---
title: "Lasso and Ridge Regression"
subtitle: "🤗"
author: "Tyson S. Barrett"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: pres.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
class: inverse, center, middle

# When is regression great?
# When is it *not* so great?

---
class: center

.pull-left[
# 🤗

.coral[
### Interpretable
### Simple
### Flexible
]
]

--

.pull-right[
# 😭

.coral[
### Correlated Predictors
### Many (but sparse) predictors
]
]

---
class: center

.pull-left[
# 🤗

.coral[
### Interpretable
### Simple
### Flexible
]
]

.pull-right[
# ~~😭~~ 🤗

.coral[
### Correlated Predictors
### Many (but sparse) predictors
]
]

<br>
<br>

.huge[**Regularized Regression** can help]

---

# Quick Review of Regression

.pull-left[
$$
Y = \beta_0 + \sum_j^p \beta_j X_j + e_i
$$

The estimated effect of $X_1$ is:
$$
\frac{\delta Y}{\delta X_1} = \beta_1
$$
]

.pull-right[

#### It finds the best fitting line to the data.
#### It can handle categorical and continuous predictors and outcomes.
#### It has decent predictive accuracy (there are better methods).
]

<br>
<br>

.center[.large[.coral[But it struggles when predictors are highly correlated and when predictor selection is necessary.]]]

---
# Regularized Regression

.pull-left[
### Built on conventional regression but includes a penalty:

$$
Y = \beta_0 + \sum_j^p \beta_j X_j + \lambda f(\beta_j) + e_i
$$
where $f(\beta_j)$ is a penalty function (i.e., if the $\beta_j$ is larger, the penalty is larger) and $\lambda$ is a tuning parameter (i.e., controls how strong the penalty is).
]

.pull-right[
### Why is this cool??
.coral[
#### It minimizes the residuals and the penalty simultaneously.
#### It gives it some magical powers.
#### It has better predictive accuracy than conventional regression.
#### Its usage is very similar to conventional regression.
]]

<br>

.center[.large[We'll discuss all of these in regards to specific techniques]]

<br>

.center[.large[But first, some general steps to using regularized regressions]]

---
background-image: url(LassoRidge_Steps.pdf)
background-size: contain
background-repeat: no-repeat

---
class: inverse, center, middle

# Popular Regularized Regressions

---
background-image: url(LassoRidge_Venn.pdf)
background-size: contain
background-repeat: no-repeat

---
# Lasso Regression

.pull-left[
### Built on conventional regression but includes a penalty:

$$
Y = \beta_0 + \sum_j^p \beta_j X_j + \lambda|\beta_j| + e_i
$$

The penalty is $|\beta_j|$ -- the absolute value of $\beta_j$.
]

.pull-right[
### This penalty makes it so:
.coral[
#### The most important variables are selected (all other have estimates equal to zero)
#### It can handle more predictors than variables (p > n).
#### It has better predictive accuracy than conventional regression. This means it is less "over-fit" and more generalizable.
#### Note: it struggles with correlated predictors.
]]

<br>

.center[.large[How do we use it?]]


---
# How do we use it?

.large[Can be done using `R` with `glmnet` or `SAS` with `proc glmselect`]

<br>

.large[I'll show it in `R` (at the end of the workshop, I'll walk through each aspect of the code)]

<br>
<br>

.large[We'll use a fake data set about two popular TV shows--The Office and Parks and Recreation.]

<br>

.coral[.large[Note: We'll be ignoring some assumptions (like the fact the data are nested).]]

---
# Dataset

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(DT)
library(tidyverse)
set.seed(84322)

## loads 'df'
load("Data/OfficeParks.RData")

df %>%
  select(1:12) %>%
  datatable(options = list(pageLength = 5))
```

---
# How do we use it?

```{r, message=FALSE, warning=FALSE, eval=FALSE}
## The R Package
library(glmnet)

## Remove Missingness (or impute)
df <- df %>%
  filter(complete.cases(prod1, depr1, marr, chil))

## Outcome
Y <- df$prod2

## Dummy Code
X <- df %>%
  select(-prod2, -nam, -start_date, -last_aired, -show)
X <- model.matrix(~ ., X)[,-1]

fit <- cv.glmnet(X, Y, alpha = 1)
plot(fit)
coef(fit, s = "lambda.min")
coef(fit, s = "lambda.1se")
```

---
# How do we use it?

.pull-left[
```{r, message=FALSE, warning=FALSE, echo=FALSE}
## The R Package
library(glmnet)

## Remove Missingness (or impute)
df <- df %>%
  filter(complete.cases(prod1, depr1, marr, chil))

## Outcome
Y <- df$prod2

## Dummy Code
X <- df %>%
  select(-prod2, -nam, -start_date, -last_aired, -show)
X <- model.matrix(~ ., X)[,-1]

fit <- cv.glmnet(X, Y, alpha = 1)
plot(fit)

```

]

.pull-right[
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(furniture)
data.frame("Lambda Min" = round(as.matrix(coef(fit, s = "lambda.min")), 5), 
           "Lambda 1SE" = round(as.matrix(coef(fit, s = "lambda.1se")), 5),
           "Variable" = row.names(coef(fit, s = "lambda.min"))) %>%
  setNames(c("Lambda.Min", "Lambda.1SE", "Variable")) %>%
  select(Variable, `Lambda.Min`, `Lambda.1SE`) %>%
  map_df(~washer(.x, 0, value = "-")) %>%
  data.frame
```
]

---
background-image: url(LassoRidge_Venn.pdf)
background-size: contain
background-repeat: no-repeat

---
# Ridge Regression

.pull-left[
### Built on conventional regression but includes a penalty:

$$
Y = \beta_0 + \sum_j^p \beta_j X_j + \lambda(\beta_j)^2 + e_i
$$

The penalty is $(\beta_j)^2$ -- the squared value of $\beta_j$.
]

.pull-right[
### This penalty makes it so:
.coral[
#### It can handle correlated predictors (even perfectly correlated)
#### It can handle far more predictors than variables (p >> n).
#### It has better predictive accuracy than conventional regression. This means it is less "over-fit" and more generalizable.
#### Note: it does not actually select variables (although the estimates can get very close to zero)
]]


.center[.large[How do we use it?]]

---
# How do we use it?

Just like Lasso but with `alpha = 0` (this will make more sense a bit later)
```{r, message=FALSE, warning=FALSE, eval=FALSE}
## The R Package
library(glmnet)

## Remove Missingness (or impute)
df <- df %>%
  filter(complete.cases(prod1, depr1, marr, chil))

## Outcome
Y <- df$prod2

## Dummy Code
X <- df %>%
  select(-prod2, -nam, -start_date, -last_aired, -show)
X <- model.matrix(~ ., X)[,-1]

fit <- cv.glmnet(X, Y, alpha = 0)
plot(fit)
coef(fit, s = "lambda.min")
coef(fit, s = "lambda.1se")
```

---
# How do we use it?

.pull-left[
```{r, message=FALSE, warning=FALSE, echo=FALSE}
## The R Package
library(glmnet)

## Remove Missingness (or impute)
df <- df %>%
  filter(complete.cases(prod1, depr1, marr, chil))

## Outcome
Y <- df$prod2

## Dummy Code
X <- df %>%
  select(-prod2, -nam, -start_date, -last_aired, -show)
X <- model.matrix(~ ., X)[,-1]

fit <- cv.glmnet(X, Y, alpha = 0)
plot(fit)
```

]

.pull-right[
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(furniture)
data.frame("Lambda Min" = round(as.matrix(coef(fit, s = "lambda.min")), 5), 
           "Lambda 1SE" = round(as.matrix(coef(fit, s = "lambda.1se")), 5),
           "Variable" = row.names(coef(fit, s = "lambda.min"))) %>%
  setNames(c("Lambda.Min", "Lambda.1SE", "Variable")) %>%
  select(Variable, `Lambda.Min`, `Lambda.1SE`) %>%
  map_df(~washer(.x, 0, value = "-")) %>%
  data.frame
```
]


---
background-image: url(LassoRidge_Venn.pdf)
background-size: contain
background-repeat: no-repeat

---
# Elastic Net

.pull-left[
#### It is the combination of Lasso and Ridge Regressions.

$$
Y = \beta_0 + \sum_j^p \beta_j X_j + \lambda\big[\alpha |\beta_j| + (1- \alpha) (\beta_j)^2\big] + e_i
$$

Note that it has a new tuning parameter, $\alpha$. Where have we seen that before?
]

.pull-right[
### This penalty makes it so:
.coral[
#### It can handle correlated predictors (even perfectly correlated)
#### It can handle far more predictors than variables (p >> n).
#### It has better predictive accuracy than conventional regression. This means it is less "over-fit" and more generalizable.
#### It selects the most important variables.
#### I.e. It combines the strengths of Lasso and Ridge.
]]


.center[.large[Again, how do we use it?]]

---
# How do we use it?

Just like Lasso but with `alpha > 0` and `alpha < 1`
```{r, message=FALSE, warning=FALSE, eval=FALSE}
## The R Package
library(glmnet)

## Remove Missingness (or impute)
df <- df %>%
  filter(complete.cases(prod1, depr1, marr, chil))

## Outcome
Y <- df$prod2

## Dummy Code
X <- df %>%
  select(-prod2, -nam, -start_date, -last_aired, -show)
X <- model.matrix(~ ., X)[,-1]

fit <- cv.glmnet(X, Y, alpha = 0.5)
plot(fit)
coef(fit, s = "lambda.min")
coef(fit, s = "lambda.1se")
```

---
# How do we use it?

.pull-left[
```{r, message=FALSE, warning=FALSE, echo=FALSE}
## The R Package
library(glmnet)

## Remove Missingness (or impute)
df <- df %>%
  filter(complete.cases(prod1, depr1, marr, chil))

## Outcome
Y <- df$prod2

## Dummy Code
X <- df %>%
  select(-prod2, -nam, -start_date, -last_aired, -show)
X <- model.matrix(~ ., X)[,-1]

fit <- cv.glmnet(X, Y, alpha = 0.5)
plot(fit)
```

]

.pull-right[
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(furniture)
data.frame("Lambda Min" = round(as.matrix(coef(fit, s = "lambda.min")), 5), 
           "Lambda 1SE" = round(as.matrix(coef(fit, s = "lambda.1se")), 5),
           "Variable" = row.names(coef(fit, s = "lambda.min"))) %>%
  setNames(c("Lambda.Min", "Lambda.1SE", "Variable")) %>%
  select(Variable, `Lambda.Min`, `Lambda.1SE`) %>%
  map_df(~washer(.x, 0, value = "-")) %>%
  data.frame
```
]

---
# Elastic Net

### How do we select the tuning parameters?

With $\lambda$ it is fairly automatic with the `cv.glmnet()` function. But what about $\alpha$?

1. Base it on theory.
2. Cross-validated error just like with $\lambda$.
3. Try a few values and see which is lowest.

<br>

None are amazing, but they all are good options.


---
# What if we want unbiased results?

### Use the best model and fit an un-penalized model.

This also gives us standard errors and p-values, which we often want for publications.

Note that this can be problemmatic if there are still a lot of selected variables or the variables are highly correlated.


---
# Other Notable Topics
.large[
1. Generalized Additive Models (GAMs)
1. Tree-based Methods
  1. Random Forests
  1. Ada Boost
  1. Fast-and-Frugal Trees
]


---




